---
title: "Lab 5"
author: "MaryJo Nelson"
date: "`r format(Sys.Date())`"
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

Welcome to my Lab 5 report. In this report, I will rename and create new variables, summarize data sets, and visualize the relationship between variables using Edgar Anderson's Iris Data, a data set with observations of 150 Iris flowers of 3 different species. 

## Load Packages and Read Data

```{r}

library(tidyverse)
(iris <- as_tibble(iris))

```

## Questions

>Question 1

Rename each variable so that it is all lower-case and uses an underscore
\_ instead of a period . in the name (the recommended coding style in
the tidyverse style guide). Print the resulting table.

```{r}

(i2 <- rename(iris,
       sepal_length = Sepal.Length,
       sepal_width = Sepal.Width,
       petal_length = Petal.Length,
       petal_width = Petal.Width,
       species = Species))

```
>Question 2

Convert the four numerical variables from cm to mm by multiplying by 10.
Print the resulting table.

```{r}

(i2_mm <- transmute(i2, 
       sepal_length_mm = sepal_length * 10,
       sepal_width_mm = sepal_width * 10,
       petal_length_mm = petal_length * 10,
       petal_width_mm = petal_width * 10,
       species = species))


```
>Question 3

Calculate sepal area and petal area (area is equal to length multiplied
by width). Print a table with only the variables sepal area, petal area,
and species.

```{r}

(areas <- transmute(i2_mm,
                      sepal_area = sepal_length_mm * sepal_width_mm,
                      petal_area = petal_length_mm * petal_width_mm,
                    species = species))

```

>Question 4

Calculate the following statistics for the entire dataset from the sepal
length variable and print the resulting table:

-   sample size

-   maximum value

-   minimum value

-   range

-   median

-   first quartile (q1)

-   third quartile (q2)

-   inter-quartile range (iqr)

```{r}

(summary_sepal_length <- summarize(i2_mm,
          sample_size = n(),
          max_value = max(sepal_length_mm),
          min_value = min(sepal_length_mm),
          range = max_value - min_value,
          median = median(sepal_length_mm),
          first_quartile = quantile(sepal_length_mm, .25),   
          third_quartile = quantile(sepal_length_mm, .75),
          inter_quartile_range = IQR(sepal_length_mm)))

```
>Question 5

Calculate the following statistics for each species from the petal width
variable and print the resulting table:

-   sample size

-   mean

-   standard deviation

-   variance

-   standard error of the mean

-   approximate 95% confidence interval

```{r}

species_grouped <- group_by(i2_mm, species)

(summary_petal_width <- summarize(species_grouped,
                                  sample_size = n(),
                                  mean = mean(petal_width_mm),
                                  sd = sd(petal_width_mm),
                                  var = var(petal_width_mm),
                                  sem = sd / sqrt(sample_size),
                                  ci_upper = mean + 2*sem,   
                                  ci_lower = mean - 2*sem))

```

>Question 6

Visualize the relationship between petal width and species using a strip
plot.

```{r}

ggplot(data = i2_mm) +
  geom_jitter(mapping = aes(y = petal_width_mm, x = species), shape = 1,
              alpha = .7) +
  labs(x = "Species", y = "Petal Width (mm)", title = "Petal widths of 150 Iris flowers organized by species (50 flowers per species)")

```

>Question 7

Starting with the previous graph, add the mean and 95% confidence
interval for each species

```{r}

ggplot(data = i2_mm) +
  geom_jitter(mapping = aes(y = petal_width_mm, x = species), shape = 1,
              alpha = .7) +
  geom_crossbar(data = summary_petal_width, mapping = aes(x = species, y = mean, 
                                                          ymax = ci_upper, ymin = ci_lower),
                color = "#EDDA74") +
  labs(x = "Species", y = "Petal Width (mm)", title = "Petal widths of 150 Iris flowers organized by species (50 flowers per species)",
       subtitle = "*Box plots show mean +/- two standard errors of the mean (roughly 95% confidence interval)")

```

>Question 8

Visualize the relationship between petal length, petal width, and
species using a scatterplot. Map the two numerical variables to the x
and y axes and map species to the color and shape aesthetics.

```{r}

ggplot(data = i2_mm) +
  geom_point(mapping = aes(x = petal_length_mm, y = petal_width_mm, color = species, shape = species),
               alpha = .5) +
  labs(x = "Petal Length (mm)", y = "Petal Width (mm)", color = "Species", shape = "Species", 
       title = "Petal width vs. petal length in 150 Iris flowers, arranged by species")

```

## Session Info

```{r}
sessioninfo::session_info()
```
